\documentclass[a4paper, 12pt]{article}

% packages
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{hyperref}


\begin{document}

\begin{titlepage}
 \title{COP507 - Computer Vision \& Embedded Systems}
 \date{29th December 2017}
 \author{ B717426}
 \maketitle 
\end{titlepage}
 
 \newpage
 
 \tableofcontents{}
 
 \newpage
 %---------------- introduction ------------ %
 
 \section{Introduction}
 \onehalfspacing
 Make and Model Recognition (VMMR) has been a subject area of interest within computer vision both in industry and academia. The importance of the ability to recognise a variety of vehicles spans across different disciplines and applications such as driver assistance, traffic monitoring, law enforcement and surveillance. As stated by Dehghan et al (2017), this categorisation task has been a problem for classical computing however, a result of great advances in artificial intelligence, this task has been made possible with accuracy rates above 90 percent. The task given was to design and implement a VMMR system with the capability to recognise the make and model of a vehicle based on only the front view (that is, headlights, grill and bumper). Based on the image of a vehicle fed as input, the output of the system should be the make and model of the vehicle. This report explains a practical use of the VMMR system that was implemented, furthermore, it explains the design of the system and any performance restrictions or performance enhancements that will be expected should the system implemented in hardware.
\parskip 0.2in 

There some basic processes that are essential for any computer vision application. These include pre-processing the acquired images, feature detection and feature extraction. Pre-processing involves making the image data conducive for feature extraction. Some of these processes includes noise reduction, morphological operations, colour correction and image re-sampling. Feature extraction is the process by which key points are obtained from the image. These key points consist of blobs, edges and corners.
 
 \newpage
 
 %--------------- Section 2: Preprocessing Tasks ------------------ % 
 \section{Preprocessing Tasks}
 This section discusses the pre-processing tasks that are required to implement the Vehicle Make and Model System in practical scenarios.
 
 \subsection{CCTV camera located at the side of a motor way}
 As a consequence of this system being implemented outdoors, there are certain factors that cannot be controlled prior to the CCTV footage being taken. These factors include illumination, motion blur, problems with focus and the camera being placed at an angle  which affects processing of the image. The following subsections explain solutions to these problems found when processing the images taken by the camera. 
 
 \subsubsection{Skewing \& Rotation}
Considering the camera is at the side of the motor way, one thing to consider is that the angle at which the camera is placed. Therefore, before performing any feature detection or extraction, the images taken from that camera must be set to the right orientation. This could be done by skewing the image or rotating the image as this will set the image to the right orientation. Traditional mathematical algorithms and formulas exist which can be used to correct the angle of rotation or the degree of skewing required. Considering that since an image can be represented as a matrix of pixel values, affine transformation matrices could be applied to the image to correct the problem (Gonzales \& Woods, 2008).
\parskip 0.2in 

Before these algorithms can be applied to the images, it must be detected whether there is a need to apply skew correction to the images captured. Patel et al. (2015) discuss an approach for skew angle detection and correction based on the radon transform. The radon transform was used because of its accuracy and low computational cost. The radon function calculates projections of the image matrix in particular directions (Patel et al, 2015). For an image matrix \textit{f(x,y)} a projection is considered as a collection of line integrals. The line integrals are calculated from various sources along beams, lateral paths, in a specific direction. The skew detection algorithm proposed by Patel et al (2015) is as follows. First of all, the input image is binarised and morphological filling operations are applied. After that, edges are detected using the canny edge detector. This process is used for locating relevant transitions within the image. The radon transform is then applied on the output of the edges detected. Next, a maximum value is searched for within the radon transform coefficients. After that, a skew angle is found from the maximum intensity value that was found from the radon coefficients. The image is then rotated with the skew angle. 

\subsubsection{ Brightness \& Colour Correction }
Another factor to consider is the lighting conditions. The reason being that, in the day time, pictures taken may or may not have the appropriate lighting. In the night time, the images taken may be very dark. Therefore, in order to make the images conducive for training, the images will need to be made brighter. A remedy to this problem could be applying a brightness adjustment algorithm. According to Szeliski (2010), this can be done by multiplying or adding the image by the constant. 

\begin{center}
 $g(x) = \alpha f(x) + \beta$
\end{center}

Where $\alpha$ and $\beta$ are parameters which influence the contrast and brightness of the image and where $\alpha$ is greater than zero. 

Alternatively, colour balancing and correction algorithms can be used to improve the illumination and contrast of the image. Weng et al (2004) introduce a novel approach of improving white balance in still images. This algorithm is needed especially for images taken at night because, depending on the colour temperature of the street light. The image will appear reddish if the colour temperature of the light source is low, conversely, the image will appear blue if the colour temperature of the light source is high (Weng et al, 2014). Therefore, improving the white balance will make the image appear as though it was captured under canonical lighting. 
\parskip 0.2in

The proposed solution works in two stages that is, white point detection and white point adjustment. utilises a dynamic threshold for identifying white points within the image. The process begins by converting the image from the $RGB$ colour space to the  $YC_{b}C_{r}$. Depending on the colour characteristics, a near white region composed of several candidate reference white points is chosen. The candidate points are chosen by computing the mean values of $C_{b}$ and $C_{r}$ (Weng et al, 2014). The white reference points are then selected depending on the brightness values  of candidate white points (top 10 percent) near the white region. After the white reference points have been chosen, the Von Kries model is used to fine-tune the image. This model is used to scale each invidivual $R, G$ and $B$ channel separately. Channel gains are obtained from the average values of reference white points. Luminance is maintained throughout the image at the same intensity by using the maximum luminance value to compute channel gains. Each pixel value is adjusted by multiplying each pixel with teh computed gain (Weng et al, 2014). 

\subsubsection{ Motion blur removal}

Motion blur is a phenomenon which occurs when there are inconsistencies between the speed of the moving object and the rate at which the camera captures frames. The problem with motion blur is that, it attenuates the image signal. Removing the motion blur from the image involves two processes. Firstly, estimating the amount by which the image has been blurred and thereafter, recovering a more realistic image using  deconvolution (point-wise division in the frequency domain). Cho (2010) proposes a method of deconvolution referred to as iterative distributed reweighting (IDR) for removing motion blur. This method enhances the visual quality of reconstructed images as compared with traditional deconvolution algorithms used for removing motion blur.
\parskip 0.2in 

Cho (2010) suggests the estimation of the blur kernel by examining edges within the blurred image. The edges within the blurred image encode estimations of the blur kernel from which the blur can be removed using the inverse radon transform. Using the edge information in this case will yield desirable results because, the object of interest is a man made object and man made objects consist of many edges. As a result of the loss of high frequency information, prior information concerning natural images is used to fill out the missing information. This is done using content aware image priors rather than sparse priors because, content aware prior adjusts its features to local textures and hence, it increases the quality of textures within the reconstructed images (Cho, 2010). After this has been completed, a maximum a posteriori (MAP) estimator is used to restore the image using piecewise smooth features, however, since the MAP does not always accurately reconstruct rich textures, IDR is used to improve the depiction of rich textures. 

\subsubsection{Shake Removal}
In the process of capturing images, the camera could be susceptible to shaking especially in windy conditions. This is a problem because, the camera shaking causes non-uniform blurring cannot be handled effectively by blind convolution algorithms (Hirsch et al, 2011). They suggest a fast single image  blind deconvolution algorithm which is a combination of both constraints of Projective Motion Path Blur (PMPB) models and the effectiveness of the Efficient Filter Flow (EFF) framework. In order to recover a sharper version of the image, EFF estimates a non-static blur as the total of various blur patches. The blur kernels are created using homographies. The homographies are then applied to a grid of pixels. This makes it possible to generate possible camera shakes via a combination of various homographies of a point grid (Hirsch et al, 2011). Different views of a point grid are created by using a homography. This splits the views into local blur kernels which are constrained to only linear combinations. 

\parskip 0.2in
Given a blurred image, an unknown sharp  image can be reconstructed in two steps: (1) blur estimation for non-stationary PSFs and (2) sharp image recovery through a non-blind deconvolution process suited to non-static blurs. The first step of the algorithm aims at retrieve the motion which the camera undertook during exposure provided only the blurry image. This involves a prediction phase, to lower the blur effect and improve image quality bilateral filtering, a blur parameter estimation phase to identify camera motion features which correctly describe the blurred image from the previous step and a latent image estimation using non-blind deconvolution (Hirsch et al, 2011). Blur parameters are approximated by finding the most appropriate non-static blur which converts the  immediate image estimate into the recorded blurry image. Sharp image updates are applied repeatedly during the blur estimation phase. After the blur parameters have been approximated and set, the final sharp image is reconstructed by replacing the image prior of the sharp image update phase with a more natural image prior which is formed on sparsity of gradient images (Hirsch et al, 2011). stationary non-blind deconvolution is then used in the non-stationary case to retrieve the sharp image. 

\newpage
\subsection{Dedicated camera installed at entrance to indoor car park}

Similar to the CCTV camera located at the side of the road, this system is affected by certain factors which may affect the quality of the images captured by the camera although not as much as an outdoor system. For instance, location at which the camera is placed can be controlled and hence it is placed at the best location possible (that is, the entrance of the car park).

\subsubsection{Noise Removal}

Noisy images might be a problem that may be experienced with images taken in camera located in the indoor car park. This noise occurs as a result of arbitrary variances in image intensities which appear as grains in the image (Kaur, 2015) . Noise is undesirable because it distorts the image signal and reduces the image quality. Kaur (2015) explains some primary sources of noise in images. These include poor light levels and the temperature of the sensor, electromagnetic transmission of the image data, effects of the environment on the image sensor, introduction of noise by the mechanism capturing image data in digital format, dust particles present on scanner screen. Noise can be removed using various techniques however the technique explained in this report is filtering (linear and non-linear). 
\parskip 0.2in

Linear filters eliminate certain types of noise by convolving the noisy image with a kernel which serves as a low pass filter (Kaur, 2015). Non-linear filtering on the other hand, produces an output which is not a linear function of its input. Using a gaussian filter to remove noise will be a viable solution however, the problem with using a gaussian filter is that, although it removes the noise, it also blurs the sharp edges and destroys useful details within the image. Therefore, using non-linear filtering is a much better solution because, non-linear filters preserve the image details in the image. For this scenario the preferred filtering technique is using a median filter. Median filters work by using a sliding window principle over the image matrix and replaces pixel values with the median value of neighbouring values. Initially, all pixels values are arranged from neighbourhood to numerical order. The pixel being considered the middle pixel value is then replaced (Kaur, 2015). Moreover, the median filter although simple, is very powerful and easy to implement. 

\subsubsection{Improving Low resolution images}

Assuming that the camera being used captures low resolution images, it is essential that the quality of the images are improved before being passed to the next phase. Similar to noise removal, there are various methods of improving the image quality. however in this scenario, Freeman et al (2002) present a method of improving image quality using super-resolution. Their approach involves using interpolation algorithms to generate tenable high frequency details in zoomed images from a database of training images. This can be used to enhance the quality of low resolution images. 
\parskip 0.2in

The algorithm works by upscaling the images using cubic spline interpolation and downscaling using convolution with a [0.25 0.5 0.25] filter and subsampling
when indicies obtained are even. The super-resolution algorithm then anticipates the next octave up that is, the lost frequencies from the image using cubic interpolation. The result the algorithm produces is total of the inputs and high frequency estimates. The high frequencies are estimated for $M$ x $N$ patches in raster scan order (Freeman et al, 2002). The algorithm functions under the presumption that, the association between high and low frequency patches is not affected by contrast, hence each patch is normalised by the median absolute value of the low frequency patches in the colour channels. These frequency patches are then merged together to form a vector that is used to search for matches in the training set. The matching is accomplished by locating the closest neighbour. When matches are located, the contrast normalisation is undone on the matching high frequency patch and that patch is added to the output image.  


\subsubsection{ Removal of Barrelling Effect}

According to Gribbon et al (2003), The barrelling effect also known as barrel distortion is a phenomenon which happens when the center of magnification of the lens is more prominent than at the boundaries (edges) of the lens.
 

\newpage

\section{Reference List} 

Cho,T.,S. (2010). \textit{Motion blur removal from photographs.} Submitted to the Department of Electrical Engineering and Computer Science in partial fulfilment of the requirements for the degree of Doctor of Philosophy in Electrical Engineering and Computer Science. Cambridge Massachusetts: Massachusetts Institute of Technology.
 
Freeman, W., J., T., Jones, T., R.,  Pasztor, E., C. (2002). Example-based super-resolution. \textit{IEEE Computer Graphics and Applications}, [Online] 22(2), pp.56-65. Available from: \url{http://ieeexplore.ieee.org/document/988747/} [Accessed 6 Jan. 2018].

Gonzalez, C., R., Woods, E., R. (2008). \textit{Digital Image Processing.} Upper Saddle River New Jersey: Pearson.

Gribbon, K. T., Johnston, C. T., and Bailey, D.
G. \textit{A Real-time FPGA Implementation of a
Lens Distortion Correction Algorithm with
Bilinear Interpolation.} Proceedings of the
Image and Vision Computing New Zealand
Conference 2003, Massey University,
Palmerston North, New Zealand, Nov. 2003. pp. 408-413,

Hirsch, M., Schuler, C., Harmeling, S. and Scholkopf, B. (2011). Fast removal of non-uniform camera shake. 2011 International Conference on Computer Vision. [online] Available from: \url{http://ieeexplore.ieee.org/document/6126276/} [Accessed 5 Jan. 2018].

Kaur, S. (2015). Noise Types and Various Removal Techniques. \textit{International Journal of Advanced Research in Electronics and Communication Engineering (IJARECE). 4(2).} [Online] p.226-229. Available from \url{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.683.6783&rep=rep1&type=pdf
} [Accessed 5th January 2018].

Patel, J., Shah, A., Patel, H. (2015). Skew Angle Detection and Correction using Radon Transform. \textit{International Journal of Electronics, Electrical and Computational System.} [Online] Volume 4 (May) Available from: \url{http://academicscience.co.in/admin/resources/project/paper/f201505101431314896.pdf}. [Accessed: 2nd January 2018].

Szeliski, R. (2010). \textit{Computer Vision: Algorithms and Applications.} New York: Springer.

Weng, C., C. Homer, C. Fuh, C., S. A Novel Automatic White Balance Method For Digital Still Cameras. 2005 \textit{IEEE International Symposium on Circuits and Systems.} [Online] Available from: \url{http://ieeexplore.ieee.org/abstract/document/1465458/} [Accessed 4 Jan. 2018].


\end{document}
